---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
activityData <- read.csv("activity.csv")
#lets create a dataframe that only includes valid step values (ie. no NA)
validData <- activityData[!is.na(activityData$steps),]
library(data.table)
DT <- data.table(validData)
aggregatedByDate <- DT[, sum(steps), by = date]
aggregatedByInterval <- DT[, sum(steps), by = interval]
setnames(aggregatedByDate,'V1','steps')
setnames(aggregatedByInterval,'V1','steps')
```


## What is mean total number of steps taken per day?
```{r}
hist(aggregatedByDate$steps)
mean(aggregatedByDate$steps)
median(aggregatedByDate$steps)
```


## What is the average daily activity pattern?
```{r}
plot(aggregatedByInterval$interval,aggregatedByInterval$steps)
````
### the interval with the most activity is `r aggregatedByInterval[aggregatedByInterval$steps==max(aggregatedByInterval$steps),interval] `

## Imputing missing values
### There is a total number of `r sum(is.na(activityData$steps))` missing values in the dataset.
### The strategy used to fill in a new values for a missing value is as follows:
1. Set the counter and sum to 0.
2. Starting with the missing value, work your way backwards until you either find a non-missing value or you reach the very begining of the list. If you find a value, add this value to the sum and increment the counter.
3. Starting with the missing value, work your way forward until you either find a non-missing value or you reach the end of the list. If you do find a value, intrement the counter and sum.
4. Take the avarage of the value(s) you have found by dividing the sum by the counter.
5. Replace the NA value with this value.
```{r}
cache=TRUE
size = nrow(activityData)
 for(i in 1:size) { 
     if(is.na(activityData[i,"steps"])) {
      sum = 0;
      counter = 0;
      j = i -1;
      while(j > 0) {
        if(!is.na(activityData[j,"steps"])) {
          sum = activityData[j,"steps"]
          counter = counter + 1
          break
        }
        j = j -1
      }
      j = i + 1
      while(j < size) {
         if(!is.na(activityData[j,"steps"])) {
          sum = activityData[j,"steps"]
          counter = counter + 1
          break
        }
         j = j + 1
      }
      average = sum/counter
      activityData[i,"steps"] <- average
     }     
  }
````

###Now that the NA values have been replaced:
```{r}
DT2 <- data.table(activityData)
aggregatedByDate2 <- DT2[, sum(steps), by = date]
aggregatedByInterval2 <- DT2[, sum(steps), by = interval]
setnames(aggregatedByDate2,'V1','steps')
hist(aggregatedByDate2$steps)
mean(aggregatedByDate2$steps)
median(aggregatedByDate2$steps)
````

## Are there differences in activity patterns between weekdays and weekends?
